---
title: "CIND820"
author: "LoD"
date: "2023-03-13"
output:
  html_document: default
  pdf_document: default
---

---
title: 'CIND820 - CLASSIFICATION ALGORITHMS - RFE'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<center> <h1> Initial Code (10%) </h1> </center>
<center>  <h3> [LoD] </h2> </center>
<center> <h3> [CIND820] </h2> </center>
---

#### In this notebook, three classification (decision tree, neural network and random forest) techniques will be applied on the RFE dataset.

#### Section 1 -  In this section, the data will be loaded and split into the testing/training sets.

#### 1. Install packages
```{r}
# install.packages("rpart.plot")
# install.packages('caret')
# install.packages("ggplot2")
```

#### 2. Load libraries
```{r}
library(caret)
library(rpart)
library(rpart.plot)
```

#### 3. Load file.
```{r}
data <- read.csv("C:/Users/-----/Desktop/Data Analytics, Big Data, and Predictive Analytics/CIND820/Data/data_optimized_RFE.csv")
```

#### 4. Split data into its training/testing sets (70 - 30 split). Set seed.
```{r}
set.seed(11)
train <- createDataPartition(data[,"Grant.Rate.Qualifier"],p=0.7,list=FALSE)
data_train <- data[train,]
data_test <- data[-train,]
```

#### 5. Convert to factors as the classification algorithms require the variables to be factors.
```{r}
data_test <- as.data.frame(lapply(data_test, as.factor))
data_train <- as.data.frame(lapply(data_train, as.factor))
```

#### 6. Implementing cross validation. 5 folds will be used.
```{r}
ctrl <- trainControl(method = "cv", number = 5)
```

####(2) 1st classification algorithm - decision tree

#### 7. build decision tree using the decoded dataset.
```{r}
tree <- train(Grant.Rate.Qualifier ~., data = data_train, method = "rpart", trControl = ctrl, tuneLength = 30)
summary(tree)
```

#### 8. predict model based on the testing data
```{r}
pred_dt <- predict(tree, data_test)
```

#### 9. confusion matrix
```{r}
cm_dt <- confusionMatrix(table(data_test[,"Grant.Rate.Qualifier"],pred_dt))
cm_dt
```

#### 10. visualize decision tree
```{r}
print(tree)
rpart.plot(tree$finalModel,extra = 106)
```

#### (3) 2nd classification algorithm - neural network

#### 11. Install packages
```{r}
# install.packages("neuralnet")
```

#### 12. Load library
```{r}
library(neuralnet)
```

#### 13. Convert to data to a matrix. Neural network uses encoded variables.
```{r}
dummy <- model.matrix( 
  ~ data_train$Grant.Rate.Qualifier + data_train$Review.Type + data_train$Decision.Type.Group + data_train$Decision.Purpose, 
  data = data_train)
head(dummy)
```

#### 14. drop first column since the conversion added an extra column.
```{r}
dummy <- dummy[,-1]
head(dummy)
```

#### 15. set as data frame
```{r}
dummy <- as.data.frame(dummy)
head(dummy)
```

#### 16. Rename column names in the dummy dataframe. 
```{r}
colnames(dummy) <- c('Grant.Rate','Review.TypePANEL','Review.TypePANEL_COMM','Review.TypePANEL_INT','Review.TypePANEL_TEL',                   'Review.TypePANEL_VID','Review.TypePAPER','Decision.Type.GroupESC','Decision.Type.GroupFULL',                   'Decision.Type.GroupUTA','Decision.PurposeBY_EXC','Decision.PurposeREG')
```

#### 17. build neural network
```{r}
net <- neuralnet(Grant.Rate ~ Review.TypePANEL + Review.TypePANEL_COMM + Review.TypePANEL_INT + Review.TypePANEL_TEL +
                   Review.TypePANEL_VID + Review.TypePAPER + Decision.Type.GroupESC + Decision.Type.GroupFULL +
                   Decision.Type.GroupUTA + Decision.PurposeBY_EXC + Decision.PurposeREG, data = dummy, hidden = 5, rep = 2, stepmax = 400000)
```

#### 18. visualize neural network
```{r}
plot(net)
```

#### 19. see neural network results
```{r}
net$result.matrix
```

#### 20. convert data_test into dummy data for prediction
```{r}
dummy2 <- model.matrix( ~ data_test$Grant.Rate.Qualifier + data_test$Review.Type + data_test$Decision.Type.Group + data_test$Decision.Purpose, 
  data = data_test)
```

#### 21. set as data frame
```{r}
dummy2 <- as.data.frame(dummy2)
```

#### 22. convert column names
```{r}
colnames(dummy2) <- c('Grant.Rate','Review.TypePANEL', 'Review.TypePANEL_COMM','Review.TypePANEL_INT','Review.TypePANEL_TEL',
                     'Review.TypePANEL_VID','Review.TypePAPER','Decision.Type.GroupESC','Decision.Type.GroupFULL',
                     'Decision.Type.GroupUTA','Decision.PurposeBY_EXC','Decision.PurposeREG')
```

#### 23. predict model based on the testing data
```{r}
pred_nn <- predict(net, dummy2)
pred_nn
```

#### (5) 3rd classification algorithm - random forest

#### 24. install libraries
```{r}
# install.packages("randomForest")
```

#### 25. load libraries
```{r}
library(randomForest)
```

#### 26. build random forest using the decoded dataset.
```{r}
rf <- train(Grant.Rate.Qualifier ~., data = data_train, method = 'rf', metric = 'Accuracy', importance = TRUE, trControl = ctrl)
```

#### 27. predict model based on the testing data
```{r}
pred_rf <- predict(rf,data_test)
```

#### 28. confusion matrix
```{r}
cm_rf <- confusionMatrix(data_test$Grant.Rate.Qualifier, pred_rf)
cm
```

#### 29. variable importance
```{r}
varImp(rf)
```
